# Summary & Overview Functions {#intro}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
devtools::install_github("kjhealy/nycdogs")
library(nycdogs)
library(ggplot2) # for mpg dataset

```

Basic dataframe for some of the examples below.

```{r}
d <- data.frame(     
  id = 1:1000,     
  x = rnorm(1000, mean = 0, sd = 1),    
  y = rnorm(1000, mean = 10, sd = 2),     
  color = sample(c('red', 'blue'), size = 1000, replace = TRUE)    
) 
```







You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].









## [RStudio Cheatsheets (master list)](https://www.rstudio.com/resources/cheatsheets/)

These seem handy.
<br><br>



Some functions that can give you a sense of the data you're working with.

List of these functions:     
- `summary`
- `glimpse`   
  
- `table`
- `dim`    
- `nrow`    
- `ncol`    
- `count`  
- `names`
<br><br>
 
## `summary`

Provides a summary of each data series in a dataframe. Good for getting a sense of the range of the
data as well as outliers and missing values. 

```{r}
summary(nyc_license)
```
`summary` behaves differently depending on the objects it's applied to:

Above we ran summary on the dataframe `nyc_license`. Here we can create a model `mod`.
We can run the `summary` function on each and get very different results.     


```{r}
mod <- lm(y ~ x, data = d)

summary(d)
summary(mod)
```
<br><br>




## `print`, `glimpse` & `str`
      
Alternative ways to see the contents of the data in more detail than `summary()` provides.     

```{r}
print(mpg)
```
Note that the data type is also included.

```{r}
glimpse(mpg)
```

```{r}
str(mpg)
```

(I think `str()` might have some powers beyond the above to unpack R objects.)
<br><br>
 

## `dim`, `row` and `col`

Shape, rows and columns of dataframe:

```{r} 
dim(nyc_license)
nrow(nyc_license)
ncol(nyc_license)
```  
<br><br>

## `count()`

Useful count of unique combinations. 
Easiest to understand with an example:

```{r}
count(mpg,manufacturer, class)
```
chevrolet has 5 different 2 seaters across the dataset.  Using `filter`, you can
see these are 5 different corvettes:
```{r}
filter(mpg,manufacturer=="chevrolet", class=="2seater" )
```
<br><br>

## `table()`

`table()` is similar to `count()`, but provides you with the counts for all possible
combinations, even if the value is 0. 


```{r}
table(mpg$manufacturer, mpg$class)
```



## `head()` and `tail()` Functions in R

```{r}

head(d, 5) # where d is dataframe

```

```{r, results="hide"}

tail(d, 5)
```

Not that by including 'results="hide" in the code chunk, the code is displayed but the results are not. 
<br><br>

## `names()`

`names` isn't a particularly useful overview function, but it can be useful to
pull column headings in a vector, which can then be referenced like any other vector. 

```{r}
names(nyc_license)

```
<br><br>

```{r}
names(nyc_license)[3]
```



<br><br>